Parallel Information Set Generation for Kriegspiel

### 2012-07-10 sbac/paper.tex
text for blueprint
Placement etc

The paper is organized as follows.  In Section~\ref{Background}, we review necessary background information, including the rules of kriegspiel and the nature of the information set generation problem.  In Section ~\ref{ParSSSE}, we describe the {\sc Charm++} Parallel State Space Search Engine (ParSSSE) platform.  We then provide our experimental results in Section~\ref{Results}.  Related work is discussed in Section~\ref{Related}.  We conclude with a discussion of future work.

%\Section{Information Set Generation}
\Section{Background: Kriegspiel}
Information Set Generation is a widely adopted operation used in game trees search. Here, we present our techniques and results in the context of a particular game - Kriegspiel. However, the techniques and framework presented in this paper is general and can be applied to other imperfect games for generating information set.
%\Section{Background: Kriegspiel}
Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
game has required three parties: two competing players and one referee.  Today,
players often compete over a computer network, with the role of the referee
being automated.  One player sees and controls only the white pieces; the
other, black.  The referee track of both sets of pieces.  The players alternate
moves as in a regular chess game, but instead of openly declaring their moves,
they secretly communicate move requests to the referee.  The referee, who knows
the location of all pieces, checks the legality of the move.

@@

\Section{Information Set Generation}
%\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figure~\ref{listing} and show a full listing and the associated pictorial progression of an actual
kriegspiel game~\cite{li94chess}.  We have adapted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{fig:9W}--\ref{fig:9B}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

The nodes in the search tree at depth $n$ correspond possible to possible sequences of $n$ plies from the start state.
There is a search space variable for each ply in the game, and the values that a variable can take on are all possible
legal moves.  Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
found to be inconsistent with the actual observation given in the input, the
subtree rooted at that node is pruned (i.e., search is discontinued along that
path).  There are types of observations that must match in order for the state
to be found: check status, pawn tries, and illegal moves available/attempted.

Figure~\ref{codelisting} shows the function at the heart of our information set generation algorithm.  The actual
observations can be viewed as global, read-only variables.  All solutions (i.e., all nodes in the information set) will
be located at the same depth in the tree: the depth that is equal to the number of plies in the observation list.
Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
for the consistency of these observations at line $7$.  If the search reaches the appropriate depth and passes these
tests, then a solution has been found and is added to the information set (line $3$).

@@

\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\label{codelisting}
\end{algorithmic}
\end{boxedminipage}
\caption{Algorithm}
\end{figure}

@@

\Section{Performance Results}\label{Results}
In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance. Next, we show that our implementation performs equally well for instances from actual games.
\subsection{Analysis using Random Instance}
The performance results reported in
this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation

@@

\begin{table}
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}
\vspace{-0.1in}
\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
\label{speedups}
\vspace{-0.3in}
\end{table}

\subsubsection{Effect of Load Balancing}
\begin{figure}[h]
\centering
\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
\vspace{-0.1in}
\caption{Effect of load balancing strategy}
\label{3schemes}
\vspace{-0.2in}
\end{figure}

In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
variation in performance across the different strategies.  We did notice that across all of our experiments, the work
stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
redistribution.  This is probably also related to the fact that we are doing an all-solutions search.

\subsubsection{Effect of Problem Size}
\begin{figure}[h]
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\vspace{-0.1in}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
\vspace{-0.2in}
\label{3depths}
\end{figure}

We next study the variation in the efficiency of the parallelization of the
algorithm with {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
number of processors increases.

@@

Next, we evaluate the effect of grainsize on the problem execution time.  The results are shown in
Tables~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
clear for this particular instance.  It is possible that there would be different behavior on different problem
instances.)  When the grainsize is too small, the overall computation is hampered by excessive parallel overhead
(communication, queueing of nodes to expand, load balancing operations).  On the other hand, when the grain size is too
large, it becomes too difficult to ensure that all processors are busy and processors are not idling with no work to do
while a small number of processors are busy doing large-sized chunks of work.

@@

\subsection{Real Game Instance}
\begin{figure}[ht]
 \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
\caption{Kriegspiel position from Virgil-David 1969 with information set size 764,209.}
\label{prob4} 
\vspace{-0.2in}
\end{figure}

\begin{figure}[ht]
 \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer Chess championship.  Information set size is 5,099,257.}
\label{prob5} 
\vspace{-0.2in}
\end{figure}

We also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs. human and computer vs. computer.  Many of these were trivial to solve in a few seconds or less on a desktop computer.  For these experiments, we used a real supercomputer - Ranger at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron processors. The nodes are interconnected using Infiniband technology in a fully-CLOS topology providing a 1GBps point-to-point bandwidth. The running times and speedups for two of the more challenging instances on Ranger are shown in Tables~\ref{prob4}--\ref{tab:prob5speedups}.   These tables show the results for runs on one to $1024$ cores.  Each column corresponds to a different parameter.  In this case, the grainsize is the number of levels of the search tree that are searched in parallel.  Up to that threshold, each node that is generated spawns a new piece of work, which the {\sc Charm++} runtime schedules to execute on one of the processing elements according to its load balancing strategy.  Above the threshold the entire subtree is searched to completion on a single processor, with no additional parallel overhead.

Note that the optimal grainsize varies for the different numbers of cores and the pattern is relatively consistent across problem instances.  The optimal grainsize increases as the number of cores increases.  This makes sense, because as this parameter grows, the number of distinct pieces of work increases.  When the number of pieces of work per processor is high, much of the overhead of parallelization is wasted.  But as the number of cores increases, the work can be efficiently broken up into smaller pieces and doled out to all the cores.  With more pieces of work, it becomes easier to balance the load.

\begin{table}[tphb]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & {\bf 177$|$1} & 178$|$0.99 & 1844$|$0.96 & 493$|$0.36 \\
4 & 71$|$2.48 & 53$|$3.34 & {\bf 48$|$3.63} & 75$|$2.35 \\
16 & 34$|$5.19 & 19 $|$9.27& {\bf 12$|$13.8} & 16$|$10.7 \\
64 & 24 $|$7.25 & 7 $|$23.6& {\bf 3.7$|$47.39} & 4.1$|$43.1 \\
256 & 24 $|$ 7.1& 4.4 $|$40.1 & 1.2 $|$145& {\bf 1.0$|$172} \\
1024 & 26 $|$6.61 & 3.8 $|$46.5& 0.64 $|$276& {\bf 0.40$|$443}
\end{tabular}
\caption{Running times in seconds and speedups (after $|$), to find full information set of size 764,209 at position 10 of Virgil-David 1969. Optimal grainsize (G) for each number of cores (PE) is shown in bold}
\vspace{-0.2in}
\label{tab:prob4}
\end{table}

\begin{table}[tphb]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & 617$|$1 & {\bf 612$|$1} & 650$|$0.95 & 3312$|$0.19 \\
4 & 213$|$2.89 & 170$|$3.63 & {\bf 158$|$3.89} & 367$|$1.68 \\
16 & 83$|$7.41 & 49$|$12.4 & {\bf 40$|$15.1} & 64$|$9.57 \\
64 & 54$|$11.3 & 19$|$32.3 & {\bf 11$|$52.9} & 14.9$|$41.4 \\
256 & 54$|$11.3 & 9.8$|$62.9 & 3.5$|$172 & {\bf 3.5$|$174} \\
1024 & 58$|$10.6 & 5.5$|$111 & 1.3$|$464 & {\bf 1.2$|$536} \\
\end{tabular}
\caption{Running times, in seconds and speedups (after $|$), to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006. Optimal grainsize for each number of cores is shown in bold.}
\vspace{-0.2in}
\label{tab:prob5}
\end{table}

Note that both problems scale well even up to 1024 processors.  We achieve a speedup of $443$ in the game between humans and up to $536$ in the game between computers.  Figures~\ref{prob4} and~\ref{prob5} show the plots of the running times for the best grainsize for each core setting.  These plots show that the running time continues to decrease as the number of cores increases.

Our results compare favorably to previous work in parallel game tree search, shown in Table~\ref{bestspeedups}.  The previous best parallelization results were for parallelizing alpha-beta pruning.  Ferguson and Korf achieved a speedup of 12 on 32 cores in the game of Othello~\cite{ferguson88distributed}.  In the game of chess, Felton and Otto showed a speedup of 100 on 256 cores~\cite{felten88highly}.  This limited scalability is due to speculative loss.  Parts of the game tree that are searched during the parallel execution would have been pruned had the algorithm run sequentially.  This is a fundamental limitation in alpha-beta search.

For partially observable games, information set generation represents a different variety of tree search that does not suffer from speculative loss.  All of the nodes that are visited in parallel would have also been visited in a sequential run.  The key challenge for parallelizing information set generation is load balancing: some parts of the tree will be pruned at a higher level than others, resulting in a wide variation in cost for each subtree.  The {\sc Charm++/ParSSSE} framework is perfectly suited to handle the load balancing issues that this problem naturally presents.

Note also that information set generation is just one subroutine needed to reason about partially observable games.  Once plausible game histories have been found, an agent may simulate some sequences of moves up to previous moves made by opponents and then recursively call the information set generation procedure from the perspective of the opponent.  This results in many parallel applications of information set generation. Furthermore, given a node in an information set, it is common for an agent to perform forward search on that node (e.g., using MCTS).  With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.
\begin{table}[tphb]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\vspace{-0.2in}
\label{bestspeedups}
\end{table}

\Section{Related Work}
An overview of the theory of imperfect information games and other basic topics in game theory can be found
in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
found in~\cite{gilpin07algorithms}.  Koller \etal, showed that a Nash
equilibrium can be found in time that is polynomial in the number of nodes in the game tree~\cite{koller94fast}.  In fact, they found
that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
feasible only for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently,
computer systems tend to utilize the concept of information sets or belief states in some form or another.  A common
strategy is to estimate the value of a move by averaging the estimated value of the associated descendant for each node
in the current information set (or a random sample from the set).  While this sort of ``averaging over clairvoyance''
{\em can} lead to very poor decisions, it is nevertheless a reasonable strategy in some games (e.g.,
Scrabble~\cite{sheppard02world}).

>>>>

The paper is organized as follows.  In Section~\ref{Background}, we review necessary background information, including the rules of kriegspiel. Next, in Section~\ref{info} we describe the nature of the information set generation problem and the algorithm used.  In Section~\ref{ParSSSE}, we describe our parallel implementation, including the {\sc Charm++} Parallel State Space Search Engine (ParSSSE) platform.  We then provide our experimental results in Section~\ref{Results}.  Related work is discussed in Section~\ref{Related}.  We conclude with a discussion of future work.

%\Section{Information Set Generation}
\Section{Background: Kriegspiel}
\label{Background}
Information Set Generation is a widely adopted operation used in game trees search. Here, we present our techniques and results in the context of a particular game - Kriegspiel. However, the techniques and framework presented in this paper is general and can be applied to other imperfect games for generating information set.
%\Section{Background: Kriegspiel}
Kriegspiel, or partially observable chess,  was invented by Henry Michael Temple in 1899.  Historically, the
game has required three parties: two competing players and one referee.  Today,
players often compete over a computer network, with the role of the referee
being automated.  One player sees and controls only the white pieces; the
other, black.  The referee track of both sets of pieces.  The players alternate
moves as in a regular chess game, but instead of openly declaring their moves,
they secretly communicate move requests to the referee.  The referee, who knows
the location of all pieces, checks the legality of the move.

@@

\Section{Information Set Generation}
\label{info}
%\subsection{Information Set Generation}
The concept of information sets is perhaps best explained through an example. 
Figure~\ref{filteredlisting},~\ref{fig:figure1} and \ref{fig:figure2} show a full listing and the associated pictorial progression of an actual
kriegspiel game~\cite{li94chess}.  We have adopted the notation used by Wolfe for Berkeley Kriegspiel.  Odd (resp. even) moves are for white (resp. black).  The actual move is shown
first followed by a list of referee announcements.  Captures announcements are prefixed with an X and give the location
of the captured piece.  Similarly, pawn tries announcements are prefixed with T.  Attempted moves that were ultimately
declared illegal are shown in a list following the `:'.  For example at move 9, (Figure~\ref{filteredlisting}),
black attempts to block a potential threat along the diagonal by moving a pawn to f6, then attempts to escape or capture
a threat at f7.  When both of these attempts fail, black can infer that his king is being threatened by a protected
bishop or queen at f7.  The attempt to move to e7 is black's way of finding out whether the threatening piece is a queen
or bishop.  When this move fails, black knows it is white's queen and ultimately retreats to d8.  Note that the failed
moves by white on turns 3 and 4 are a result of being blocked by the pawn at e2.

@@

The nodes in the search tree at depth $n$ correspond possible to possible sequences of $n$ plies from the start state.
There is a search space variable for each ply in the game, and the values that a variable can take on are all possible
legal moves.  Information set generation, then, is an all-solutions search through the space.  Whenever a node in the search tree is
found to be inconsistent with the actual observation given in the input, the
subtree rooted at that node is pruned (i.e., search is discontinued along that
path).  There are types of observations that must match in order for the state
to be found: check status, pawn tries, and illegal moves available/attempted.
\begin{figure}[htpb]
\small
%\begin{algorithm}
\begin{boxedminipage}{\columnwidth}
\begin{algorithmic}[1]
\Function{DFS-Infoset}{$a_{1:t}$, $o_{1:T}^{(k)}$, $I$, $s$, $t$}
  \If {$t > T$}
    \State {$I \leftarrow I \cup a_{1:t}$} %\Comment{$a_{1:t}$ leads to $o_{1:T}$ for $P_k$; add it to the information set.}
  \Else {
    \For{\textbf{each} legal action $a$ at $s$}
      \State {$s'$ = {\sc GetNextState}($s$,$a$)}\label{line:update}
      \If {$s'.o_{t+1}^{(k)}=o_{t+1}^{(k)}$}
        \State {{\sc DFS-Infoset}($a_{1:t}a$, $o_{1:T}^{(k)}$, $I$, $s'$, $t+1$)} 
          %\Comment{ Recall that $N(a_{1:t}a) = T(T(a_{1:t}),a)$}
      \EndIf
    \EndFor
  }\EndIf
\EndFunction
\label{codelisting1}
\end{algorithmic}
\end{boxedminipage}
\caption{Algorithm}
\end{figure}

Figure~\ref{codelisting1} shows the function at the heart of our information set generation algorithm.  The actual
observations can be viewed as global, read-only variables.  All solutions (i.e., all nodes in the information set) will
be located at the same depth in the tree: the depth that is equal to the number of plies in the observation list.
Referee announcements regarding pawn tries or a player being in check are made prior to a move.  The algorithm checks
for the consistency of these observations at line $7$.  If the search reaches the appropriate depth and passes these
tests, then a solution has been found and is added to the information set (line $3$).

@@

@@

\Section{Performance Results}\label{Results}

\begin{table}[ht]
\centering
\begin{tabular}{rrr}
{\bf Procs}	&	{\bf Time (s)} 	&	{\bf Speedup}\\
1	&	1524	&	1\\
2	&	812	&	1.88\\
4	&	407	&	3.74\\
8	&	208	&	7.32\\
16	&	102	&	14.8\\
32	&	53.2	&	28.7\\
64	&	28.0	&	54.5\\
128	&	14.9	&	102\\
256	&	7.70	&	197\\
512	&	4.59	&	331\\
1024	&	2.64	&	576\\
\end{tabular}
\vspace{-0.1in}
\caption{Speedups for a randomly selected problem instance on 1--1024 processors.  The input observations were from the
10th ply of the game tree and the size of the information set in this instance is 4,495,121.}
\label{speedups}
\vspace{-0.3in}
\end{table}
In order to understand the effect of various parameters such as grain size and load balancing, we first evaluated the scalability of our implementation using a randomly selected problem instance. Next, we show that our implementation performs equally well for instances from actual games.
\subsection{Analysis using Random Instance}
The performance results reported in
this section were obtained by running our program on Blueprint with up to 1024 processors. TODO: Blueprint H/w or citation

@@

\subsubsection{Effect of Load Balancing}

In our next round of experiments, we set out to compare a few different load-balancing strategies for the problem.  We
used the random, work stealing, and neighborhood balancing strategies that are available in the general purpose search
engine.  These results are shown in Figure~\ref{3schemes}.  Our results show that there was actually very little
variation in performance across the different strategies.  We did notice that across all of our experiments, the work
stealing approach appeared to be consistently---if only marginally---better than the others.  Our hypothesis here is
that perhaps the nodes are rarely starved for work.  In other words, perhaps the very nature of the problem means that
the load is naturally well-balanced and that the load-balancer rarely needs to do a significant amount of
redistribution.  This is probably also related to the fact that we are doing an all-solutions search. 
\begin{figure}[ht]
\centering
\includegraphics[width=0.9\columnwidth]{plots/3schemes.pdf}
\vspace{-0.1in}
\caption{Effect of load balancing strategy}
\label{3schemes}
\vspace{-0.2in}
\end{figure}

@@

\begin{figure}[ht]
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depths.pdf}
\end{minipage}
\centering
\begin{minipage}{0.7\linewidth}
\includegraphics[width=\columnwidth]{plots/3depthsnolog.pdf}
\end{minipage}
\vspace{-0.1in}
\caption{(Top) Log-log plot of running time for three different sizes of problem as the number of cores used increases.
(Bottom) Same data with a linear y-axis for the running time (to help see the Iso-efficiency.)}
\vspace{-0.2in}
\label{3depths}
\end{figure}
Next, we evaluate the effect of grainsize on the problem execution time.  The results are shown in
Figures~\ref{dep8}--~\ref{dep12}.  It is clear that there is a ``sweet spot'' for each problem size.  (At least it is
clear for this particular instance.  It is possible that there would be different behavior on different problem
instances.)  When the grainsize is too small, the overall computation is hampered by excessive parallel overhead
(communication, queueing of nodes to expand, load balancing operations).  On the other hand, when the grain size is too
large, it becomes too difficult to ensure that all processors are busy and processors are not idling with no work to do
while a small number of processors are busy doing large-sized chunks of work.

@@

\subsubsection{Effect of Problem Size}

We next study the variation in the efficiency of the parallelization of the
algorithm with {\em size} of the problem?  To test this, we varied the number of plies in the test instance from
8 to 12. We used the work stealing load balancing strategy.  The results are shown in Figure~\ref{3depths}.   In the easiest case (depth 8), there isn't
enough additional parallel work to justify the jump from 256 to 1024 processors, and performance actually degrades.
Other than that, however, we see consistent speedups as we increase the number of processors.  The plot on the right helps to
illustrate the good isoefficiency of the problem.  As the size of the problem increases, the amount of parallelizable
work available also increases, and therefore the steepness of the speed curve is higher for larger problem sizes, as the
number of processors increases.
\subsection{Real Game Instance}
%\begin{figure}[ht]
% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem4.pdf} %
%\caption{Kriegspiel position from Virgil-David 1969 with information set size 764,209.}
%\label{prob4} 
%\vspace{-0.2in}
%\end{figure}

%\begin{figure}[ht]
% \centering \includegraphics[viewport= 280 200 300 550, scale=0.35]{images/KriegspielProblem5.pdf} %
%\caption{Kriegspiel position from Game 6 of Kbot-Darkboard at the 2006 Computer Chess championship.  Information set size is 5,099,257.}
%\label{prob5} 
%\vspace{-0.2in}
%\end{figure}

\begin{figure}[ht]
 \centering \includegraphics[scale=1]{plots/new.pdf} %
\caption{Kriegspiel position from 2 games, postiion 10}
\label{prob4} 
\vspace{-0.2in}
\end{figure}

\begin{table}[ht]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & {\bf 177$|$1} & 178$|$0.99 & 1844$|$0.96 & 493$|$0.36 \\
4 & 71$|$2.48 & 53$|$3.34 & {\bf 48$|$3.63} & 75$|$2.35 \\
16 & 34$|$5.19 & 19 $|$9.27& {\bf 12$|$13.8} & 16$|$10.7 \\
64 & 24 $|$7.25 & 7 $|$23.6& {\bf 3.7$|$47.39} & 4.1$|$43.1 \\
256 & 24 $|$ 7.1& 4.4 $|$40.1 & 1.2 $|$145& {\bf 1.0$|$172} \\
1024 & 26 $|$6.61 & 3.8 $|$46.5& 0.64 $|$276& {\bf 0.40$|$443}
\end{tabular}
\caption{Running times in seconds and speedups (after $|$), to find full information set of size 764,209 at position 10 of Virgil-David 1969. Optimal grainsize (G) for each number of cores (PE) is shown in bold}
\vspace{-0.2in}
\label{tab:prob4}
\end{table}

\begin{table}[th]
\centering
\begin{tabular}{rrrrr}
PE/G & 2 & 4 & 6 & 8 \\
1 & 617$|$1 & {\bf 612$|$1} & 650$|$0.95 & 3312$|$0.19 \\
4 & 213$|$2.89 & 170$|$3.63 & {\bf 158$|$3.89} & 367$|$1.68 \\
16 & 83$|$7.41 & 49$|$12.4 & {\bf 40$|$15.1} & 64$|$9.57 \\
64 & 54$|$11.3 & 19$|$32.3 & {\bf 11$|$52.9} & 14.9$|$41.4 \\
256 & 54$|$11.3 & 9.8$|$62.9 & 3.5$|$172 & {\bf 3.5$|$174} \\
1024 & 58$|$10.6 & 5.5$|$111 & 1.3$|$464 & {\bf 1.2$|$536} \\
\end{tabular}
\caption{Running times, in seconds and speedups (after $|$), to find full information set of size 5,099,257 at position 10 of Game 6 Kbot-Darkboard 2006. Optimal grainsize for each number of cores is shown in bold.}
\vspace{-0.2in}
\label{tab:prob5}
\end{table}

We also evaluated our information set generation routine on 40 positions from actual kriespiel games, including human vs. human and computer vs. computer.  Many of these were trivial to solve in a few seconds or less on a desktop computer.  For these experiments, we used a real supercomputer - Ranger at TACC~\cite{ranger}. Ranger has 16-way SMP nodes containing AMD Opteron processors. The nodes are interconnected using Infiniband technology in a fully-CLOS topology providing a 1GBps point-to-point bandwidth. The running times and speedups for two of the more challenging instances on Ranger are shown in Tables~\ref{tab:prob4}--\ref{tab:prob5}.   These tables show the results for runs on one to $1024$ cores.  Each column corresponds to a different parameter.  In this case, the grainsize is the number of levels of the search tree that are searched in parallel.  Up to that threshold, each node that is generated spawns a new piece of work, which the {\sc Charm++} runtime schedules to execute on one of the processing elements according to its load balancing strategy.  Above the threshold the entire subtree is searched to completion on a single processor, with no additional parallel overhead.

Note that the optimal grainsize varies for the different numbers of cores and the pattern is relatively consistent across problem instances.  The optimal grainsize increases as the number of cores increases.  This makes sense, because as this parameter grows, the number of distinct pieces of work increases.  When the number of pieces of work per processor is high, much of the overhead of parallelization is wasted.  But as the number of cores increases, the work can be efficiently broken up into smaller pieces and doled out to all the cores.  With more pieces of work, it becomes easier to balance the load.
Note that both problems scale well even up to 1024 processors.  We achieve a speedup of $443$ in the game between humans and up to $536$ in the game between computers.  Figures~\ref{prob4} shows the plots of the running times for the best grainsize for each core setting.  These plots show that the running time continues to decrease as the number of cores increases.

Our results compare favorably to previous work in parallel game tree search, shown in Table~\ref{bestspeedups}.  The previous best parallelization results were for parallelizing alpha-beta pruning.  Ferguson and Korf achieved a speedup of 12 on 32 cores in the game of Othello~\cite{ferguson88distributed}.  In the game of chess, Felton and Otto showed a speedup of 100 on 256 cores~\cite{felten88highly}.  This limited scalability is due to speculative loss.  Parts of the game tree that are searched during the parallel execution would have been pruned had the algorithm run sequentially.  This is a fundamental limitation in alpha-beta search. 
\begin{table}[ht]
\centering
\begin{tabular}{lrrc}
{\bf Authors} & {\bf Game} & {\bf Year} & {\bf Best Speedup (cores)} \\
Ferguson \& Korf & Othello & 1988 & 12(32) \\
Felton \& Otto & Chess & 1988 & 100(256) \\
Richards \etal & Kriegspiel & 2012 & 178(256) \\
Richards \etal & Kriegspiel & 2012 & {\bf 536(1024)} \\
\end{tabular}
\caption{Best Speedups reported for game tree applications.}
\vspace{-0.2in}
\label{bestspeedups}
\end{table}
For partially observable games, information set generation represents a different variety of tree search that does not suffer from speculative loss.  All of the nodes that are visited in parallel would have also been visited in a sequential run.  The key challenge for parallelizing information set generation is load balancing: some parts of the tree will be pruned at a higher level than others, resulting in a wide variation in cost for each subtree.  The {\sc Charm++/ParSSSE} framework is perfectly suited to handle the load balancing issues that this problem naturally presents.

Note also that information set generation is just one subroutine needed to reason about partially observable games.  Once plausible game histories have been found, an agent may simulate some sequences of moves up to previous moves made by opponents and then recursively call the information set generation procedure from the perspective of the opponent.  This results in many parallel applications of information set generation. Furthermore, given a node in an information set, it is common for an agent to perform forward search on that node (e.g., using MCTS).  With these two additional levels of parallelism on top of the information set generation problem, it is conceivable that the problem of playing partially observable games could scale efficiently to millions of processors or more.

\Section{Related Work}
\label{Related}
An overview of the theory of imperfect information games and other basic topics in game theory can be found
in~\cite{kuhn03lectures} and~\cite{kuhn97classics}.  A technical explanation of the concept of information sets can be
found in~\cite{gilpin07algorithms}.  Koller \etal, showed that a Nash
equilibrium can be found in time that is polynomial in the number of nodes in the game tree~\cite{koller94fast}.  In fact, they found
that it took about as much time to generate the full game tree as it did to actually solve it.  Unfortunately, this is
feasible only for small games.  In larger games where theoretically optimal strategies cannot be computed efficiently,
computer systems tend to utilize the concept of information sets or belief states in some form or another.  A common
strategy is to estimate the value of a move by averaging the estimated value of the associated descendant for each node
in the current information set (or a random sample from the set).  While this sort of ``averaging over clairvoyance''
{\em can} lead to very poor decisions, it is nevertheless a reasonable strategy in some games (e.g.,
Scrabble~\cite{sheppard02world}).
